### 2018-03-03 cover_letter.tex
coverletter 和 对修改回信的latex文件，添加杨涛签名

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Long Lined Cover Letter
% LaTeX Template
% Version 1.0 (1/6/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Matthew J. Miller
% http://www.matthewjmiller.net/howtos/customized-cover-letter-scripts/
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[10pt,stdletter,dateno,sigleft]{newlfm} % Extra options: 'sigleft' for a left-aligned signature, 'stdletternofrom' to remove the from address, 'letterpaper' for US letter paper - consult the newlfm class manual for more options

\usepackage{charter} % Use the Charter font for the document text
\usepackage{changepage}
\usepackage{color}

\usepackage{ amssymb }

%\newsavebox{\Luiuc}\sbox{\Luiuc}{\parbox[b]{1.75in}{\vspace{0.5in}
%\includegraphics[width=1.2\linewidth]{logo.png}}} % Company/institution logo at the top left of the page
%\makeletterhead{Uiuc}{\Lheader{\usebox{\Luiuc}}}

%\newlfmP{sigsize=50pt} % Slightly decrease the height of the signature field
\newlfmP{addrfromphone} % Print a phone number under the sender's address
\newlfmP{addrfromemail} % Print an email address under the sender's address
\PhrPhone{Phone} % Customize the "Telephone" text
\PhrEmail{Email} % Customize the "E-mail" text

%\lthUiuc % Print the company/institution logo

%----------------------------------------------------------------------------------------
%	YOUR NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------

 % Name
\signature{\\\includegraphics[width=4cm]{1.png}\\Tao Yang}
\addrfrom{
\today\\[12pt] % Date
1037 Luoyu Road \\ % Address
Wuhan, Hubei, China, 430074
}
\phonefrom{+86 15032905712}
\emailfrom{hust_yt@hust.edu.cn} % Email address

%----------------------------------------------------------------------------------------
%	ADDRESSEE AND GREETING/CLOSING
%----------------------------------------------------------------------------------------
\greetto{Dear editor,} % Greeting text
\closeline{Sincerely yours,} % Closing text

\nameto{$\langle$ Applied Energy$\rangle$} % Addressee of the letter above the to address

\addrto{
$\langle$ A novel 1000\,MW double-reheat ultra-supercritical system with \\
turbine-extraction-heated air preheaters and low-temperature economizers$\rangle$\\ % To address
}

%----------------------------------------------------------------------------------------

\begin{document}
\begin{newlfm}

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------

I enclose a manuscript entitled `A novel 1000 MW double-reheat ultra-supercritical system with turbine-extraction-heated air preheaters and low-temperature economizers'. 
All of the authors agree to the submission of this paper.

This paper belongs to the field of energy system analysis and optimization using exergy analysis method.

The manuscript includes 25 pages, 8 tables, and 10 figures prepared using Latex. 
The latex source files have been upload in a zip file.

We respectfully submit that the following individuals would be suitable peer reviewers based on their expertise in the field:
\\
\\
Inmaculada Arauzo, Universidad de Zaragoza (Spain), E-mail: iarauzo@unizar.es

Danmei Xie, Wuhan University (China), E-mail: dmxie@whu.edu.cn

Daotong Chong, Xi'an Jiao tong University (China),E-mail: dtchong@mail.xjtu.edu.cn
\\
\\

I confirm that I have mentioned all the organizations that found my research in the acknowledgment section of my submission, including grant numbers.
And the manuscript has not been published before nor submitted to another journal for the consideration of publication.

%----------------------------------------------------------------------------------------

\end{newlfm}
\end{document}